DATA 606_First_Assignment

Md_Jalal_Uddin

August 30, 2016

Problem Set 1:

1. Calculate the dot product u.v where u = [0.5, 0.5] and v = [3, -4]

```{r}

 u <- c(0.5,0.5)
 v <- c(3,-4)
 dot_product_uv<- u%*%v
 print (dot_product_uv)
    
```

2. What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

```{r}
len_u <- sqrt(sum(u*u))
 print(len_u)

len_v <- sqrt(sum(v*v))
 print (len_v)

```

 3. What is the linear combination: 3u - 2v? 
 
```{r}
lin_com <- (3*u)-(2*v)
 print (lin_com)

```

4. What is the angle between u and v

```{r}
 theta <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)) ) )  
 print(theta)
 #the angle given in radius 
#http://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r


```

Problem Set 2:



Please test it with the system below and it should produce a solution x=[???1.55;???0.32;0.95]x=[???1.55;???0.32;0.95]
?????????12???11???1???2354??????????????????x1x2x3?????????=?????????126?????????

Solution:

```{r}
# writing matrix A 

A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
A # to see the matrix A
```

 
```{r}
B <- c(1, 2, 6)
B # to see given vector

```


```{r}
# creating function to solve system of linear equation using back-substitution


# the desire linear equation

linear_equ(A,B)

#Source: 
# http://stackoverflow.com/questions/8145694/solving-simultaneous-equations-with-r
#http://www.tutor-homework.com/Math_Help/college_algebra/m5l4notes.pdf
```
