DATA 606_First_Assignment

Md_Jalal_Uddin

August 30, 2016

Problem Set 1:

1. Calculate the dot product u.v where u = [0.5, 0.5] and v = [3, -4]

```{r}

 u <- c(0.5,0.5)
 v <- c(3,-4)
 dot_product_uv<- u%*%v
 print (dot_product_uv)
    
```

2. What are the lengths of u and v? Please note that the mathematical notion of the length of a vector is not the same as a computer science definition.

```{r}
len_u <- sqrt(sum(u*u))
 print(len_u)

len_v <- sqrt(sum(v*v))
 print (len_v)

```

 3. What is the linear combination: 3u - 2v? 
 
```{r}
lin_com <- (3*u)-(2*v)
 print (lin_com)

```

4. What is the angle between u and v

```{r}
 theta <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)) ) )  
 print(theta)
 #the angle given in radius 
#http://stackoverflow.com/questions/1897704/angle-between-two-vectors-in-r


```

Problem Set 2:

Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. Please note that you do have to worry about zero pivots, and that you should not use the built-in function solve to solve this system or use matrix inverses.The approach that you should employ is to construct an Upper Triangular Matrix and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.

Please test it with the system below and it should produce a solution x=[???1.55;???0.32;0.95]x=[???1.55;???0.32;0.95]
?????????12???11???1???2354??????????????????x1x2x3?????????=?????????126?????????

Solution:

```{r}
# writing matrix A 

A <- array(c(1, 2, -1, 1, -1, -2, 3, 5, 4), dim=c(3,3))
A # to see the matrix A
```

 
```{r}
B <- c(1, 2, 6)
B # to see given vector

```


```{r}
# creating function to solve system of linear equation using back-substitution

linear_equ <- function(A, B) {
  # creating a vector.
  x = numeric(3) 
  # bind A and b to create a new matrix Y 
  Y = cbind(A, B)
  
  # need 1 in top left pivot
  if (Y[1,1] == 0) {
    #  need to switch with row that has non-zero in first column
    if (Y[2,2] != 0) {
    # switch rows 1 and 2 as row 2 has a non-zero in column 1
    Y = Y[c(2,1,3),]
    } else {
    # switch rows 3 and 1 since row 3 has a non-zero
      Y = Y[c(3,2,1),]
    }
  }  else if (Y[1,1] != 1) {
    # else divide 1st row by row 1 column 1 to get 1 in pivot
    Y[1,] <- Y[1,] / Y[1,1]
  }  

  # Zero other entries in column 1
  if (Y[2,1] !=0) {
    # use value of row 2 column 1 for vector addition to zero the row
    vec <- Y[2,1] * Y[1,]
    # reduce row 2 using the new vec object
    Y[2,] <- -Y[2,] + vec
  }
  if (Y[3,1] !=0) {
    # use value of row 3 column 1 for vector addition to zero second row
    vec <- Y[3,1] * Y[1,]
    # now reduce row 2 using the vec
    Y[3,] <- -Y[3,] + vec
  }
  
  # check if row 2 column 2 are 0. If no, switch for row 3 to get non-zero
  if (Y[2,2] == 0) {
    # switch rows 2 and 3 as row 2 has 0 in column 2
    Y = Y[c(1,3,2),]
  } 
  
  # divide second row by row 2 column 2 to get 1 in pivot
  Y[2,] <- Y[2,] / Y[2,2]

  # use value of row 3 column 2 for vector addition to zero row 3 column 2
  vec <- Y[3,2] * Y[2,]
  # now reduce row 3 using vec
  Y[3,] <- -Y[3,] + vec  

  # solve for x with back subsitution
  x[3] <- Y[3,4] / Y[3,3]
  x[2] <- (Y[2,4] - Y[2,3]*x[3]) / Y[2,2]
  x[1] <- (Y[1,4] - Y[1,2]*x[2] - Y[1,3]*x[3]) / Y[1,1]

  return(x)
}

# the desire linear equation

linear_equ(A,B)

#Source: 
# http://stackoverflow.com/questions/8145694/solving-simultaneous-equations-with-r
#http://www.tutor-homework.com/Math_Help/college_algebra/m5l4notes.pdf
```
